---
title: "MT500 Quarto Report by Dev Jagasia"
output:
  html_document:
    css: styles.css
    toc: true
    toc_depth: 3
    toc_float: false
    code_folding: show
    number_sections: true
    theme: united
    highlight: tango
---

***Code 1 generates a world map with regions shaded based on their HIV test results. To present the results, the UNICEF HIV data is combined with the world map data and plotted using ggplot2's geom_polygon function. Higher rates of HIV are represented by color schemes that are darker, whereas lower rates are represented by lighter colors***.

**Code 1- A World Map Chart**

```{r}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
install.packages(c("ggplot2", "tidyverse", "dplyr", "stringr", "maps", "mapdata", "sf", "units", "gghighlight"))
library(tidyverse)
library(sf)
unicef <- read_csv("region_sex_value_time.csv")
mapdata <- map_data("world")
mapdata <- left_join(mapdata, unicef, by = c("region" = "country_region"), suffix = c(".map", ".unicef"))
mapdata <- filter(mapdata, !is.na(alpha_2_code))
ggplot(mapdata, aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = obs_value), color = "red") + 
  labs(fill = "Value") + 
  theme(panel.background = element_rect(fill = "lightblue", color = "lightblue"), 
        panel.grid.major = element_line(color = "lightblue"))
```

***Code 2 generates a stacked bar graph that displays the HIV test results by sex over time. The graph depicts the percentage of people who reported having an HIV test over time. The bars are stacked by gender to illustrate the proportion of men and women who reported an HIV test over the time period.***

```{r}
#stacked bar graph

data <- read.csv("/Users/devjagasia/region_sex_value_time.csv")

# Create the stacked bar graph
ggplot(data, aes(x = time_period, y = obs_value, fill = sex)) +
  geom_bar(stat = "identity") +
  labs(x = "Observed Value in %", y = "Time Period", title = "HIV test reported over the years globally")

```

***Code 3 generates a scatter plot displaying the observed value of HIV test results by location and gender. The x-axis represents the alpha code region, and the y-axis represents the observed value of the HIV test results. A linear regression line and points colored by sex are included in the plot.***

```{r}
# for scatter plot

library(ggplot2)

# Load the data from the CSV file
data <- read.csv("/Users/devjagasia/region_sex_value_time.csv")

# Create the scatter plot
ggplot(data, aes(x=alpha_2_code, y=obs_value, group = sex)) +
  geom_point() +
  
  # Add a linear regression line
  geom_smooth(method="lm", se=FALSE, color = "red") +
  
  labs(x = " Region by Alpha Code ", y = " Observed Value") +
  scale_color_discrete(name = "Sex")

```

***Code 4 generates a time series visualization of HIV test results over time. The x-axis represents the year, while the y-axis represents the observed value of HIV test results. The graphic comprises a data line graph with blue colour representing HIV test results over time.***

```{r}
# for time series plot

library(ggplot2)
library(scales)

# Read in the data
data <- read.csv("/Users/devjagasia/region_sex_value_time.csv")
data$date <- as.Date(paste(data$year, "-01-01"), format = "%Y-%m-%d")

# Convert the date column to a date object
data$date <- as.Date(data$date)


ggplot(data, aes(x = year, y = obs_value)) +
  geom_line(color = "blue") +
  labs(title = "Time Series Plot", x = "Year", y = "Observed Value")


```

```{r}

```
